\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}

### 2014-06-13 paper/subdocuments/data_and_methods.Rnw
minor changes to data and methods and basic structure of results section

The central income reported through tax statistics is the taxable income. It includes all reported incomes (income from employment, income from property and received transfers\footnote{Realized capital gains are not taxed and Mean-tested benefits are not taxed and therefore not included in Tax data. Income for low income groups are therefore underestimated. However, Piketty and Saez (2001) note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had nonly a trivial impact on top income shares.}) minus several deductions. It is therefore neither a pre-transfer income  nor a post-transfer income measure. It's rather something in between. As the FTA tax statistics include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions. This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction, Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income". As some deductions can be interpreted as compulsory expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption. Similar when calculating the disposable income out of the taxable income through accounting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income). Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches, which represent the bulk of taxes in Switzerland, are missing.   \\

@@

@@

In tax data, however, the units are represented according to administrative rules. Tax units therefore neither represent individuals in every case nor true households. Tax units rather represent individuals and couples, but couples, who are married or officially registered. This doesn't imply, that those couples live together, as it is needed to satisfy the definition of a household. On the other hand, is it quit likely that more than one tax units live in the same household (unmarried/unregistered couples, see \citet[99]{muller_vermogenslage_2014}). It is therefore not directly possible to elicit households and household income out of tax data. This might influence the assessment inequality, taking into account the change from traditional household and family structures over the last century. This is an inequality related issue, where relevant studies are missing.

% Allenfalls die Br?lhart-Daten erw?hnen und die m?glichkeit der Haushaltsbildung ?ber Abz?ge. Falls wir auch eine Grafik zur Rolle der Haushaltszusammensetzung machen (ich denke schon)
% gibt es studien? Zur Bedeutung von ?quivalenzskalen und zur Bedeutung der Differenzierung von Haushalten und Individuuen?

@@

The recent wave of tax-data studies do not report Gini-coefficents. Rather top income shares are informed on, which are calculated not only with taxdata, but togehter with external sources to produce the population and personal income control totals. This procedure ensures, that the inequality measure is not biased because of non-fillers, who doesn't appear in tax statistics.  Leight (2002:594f) compares top income shares with other inequality measures and ask, if they are a useful measure of inequality in a society. He tries to answer this question empiricaly by comparing measures of inequality based on top income shares with measures of household or familiy inequality. He finds a strong positive relationship, but concludes (P.600): top income shares are far from perfect as a measure of distribution of income across society. Top income shares hence inform not completely on how inequality evolves elsewhere in the distribution. Furthermore, top income shares only weakly satisfy the Pigou-Dalton transfer principle (in contrast to the Gini-Coefficent see above). A transfer from rich to poor will indeed never increase the top income shares, but if the transfer is between invidviduals, who are either both within the top group or both outsie the top group, then the share measure will remain unchanged. \\

@@

An other critical issue is the problem of tax evasion, which definitely can bias the assessment of inequality. Alvaredo and Saez (2006) for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion. Evasion can occured, when individuals try not to fill tax returns or by misreporting of incomes. In Switzerland non-fillers show up in the tax-statistics either way, as long as they are registered. This person get an imputed income form an older tax return and infomration given by employers. Only non registered non-fillers are not in the records. Therefore non-fillers are a minor problem. Not neglectable is the circumstance, that individuals misreport incomes. \citet{feld_tax_2006} examine the role of tax evasion in Switzerland by calculating the difference of the national accounts measures of primary income and the income reported to the tax authorities. Building on a time series of Pommerehne and Weck-Hannemann (1996) Feld and Frey (2002) and Frey and Feld (2002) extend the series. They can show, that the average level of income tax evasion from 1965 to 1995 varies between 13\% and 35\%. They suggest, that evasion is heavily driven by capital income tax evasion. \\

@@

%[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
% Mögliche Vergleichsdimensionen
%- Suitability of cross country comparison
%- Coverage of the population of interesst
%- Implementation of common concepts of economic ressources
%- intertemporal comparision
%- flexibel applications of appropriate inequality measures

@@

%%
% Population coverage
% Analysen zu den Nullern aufzeigen> ist eine Hauptkritik an der Nutzung von Steuerdaten
% Die Bedeutung von Normal- und Sonderfällen
% Den Vergleich mit der HABE beschreiben, welche Einkommensgrösse nutzen wir aus der HABE um diese möglichst mit den Steuerdaten vergleichbar zu machen? Wir brauchen das Bruttoeinkommen nach Abzug der Sozialversicherungsbeiträge, dies lässt sich mit dem Reineinkommen vergleichen

>>>>

The central income reported through tax statistics is the taxable income. It includes all reported incomes (income from employment, income from property and received transfers\footnote{ Mean-tested benefits are not taxed and therefore not included in Tax data. Income for low income groups are therefore underestimated. However, Piketty and Saez (2001) note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions. It is therefore neither a pre-transfer income  nor a post-transfer income measure. It's rather something in between. As the FTA tax statistics include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions. This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction, Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income". As some deductions can be interpreted as compulsory expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption. Similar when calculating the disposable income out of the taxable income through accounting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income). Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches, which represent the bulk of taxes in Switzerland, are missing.   \\

@@

% Allenfalls darauf hinweisen, dass die Wahl der Einkommensdefinition von der Forschungsfrage abhängt. Inequality in market outcome (genderdifferences) requieres to look at market earings from employment, to assess the role of redistribution different steps of income has to be looked at. Following the idea of economic wellbeing usual disposable income is looked at

@@

In tax data, however, the units are represented according to administrative rules. Tax units therefore neither represent individuals in every case nor true households. Tax units rather represent individuals and couples, but couples, who are married or officially registered. This doesn't imply, that those couples live together, as it is needed to satisfy the definition of a household. On the other hand, is it quit likely that more than one tax units live in the same household (unmarried/unregistered couples, see \citet[99]{muller_vermogenslage_2014}). It is therefore not directly possible to elicit households and household income out of tax data. This might influence the assessment inequality, taking into account the change from traditional household and family structures over the last century.

% Allenfalls die Br?lhart-Daten erw?hnen und die m?glichkeit der Haushaltsbildung ?ber Abz?ge. Falls wir auch eine Grafik zur Rolle der Haushaltszusammensetzung machen (ich denke schon)
% gibt es studien? Zur Bedeutung von ?quivalenzskalen und zur Bedeutung der Differenzierung von Haushalten und Individuuen?
% inequality and family structure> review ny (McCall & Percheski (2010))
% Was ist zu erwarten, wenn Äquivalenzskalen vernachlässigt werden? Soll man dies erwähnen oder einfach weglassen

@@

The recent wave of tax-data studies do not report Gini-coefficents. Rather top income shares are informed on, which are calculated not only with taxdata, but togehter with external sources to produce the population and personal income control totals. This procedure ensures, that the inequality measure is not biased because of non-fillers, who doesn't appear in tax statistics.  Leight (2002:594f) compares top income shares with other inequality measures and asks, if they are a useful measure of inequality in a society. He tries to answer this question empiricaly by comparing measures of inequality based on top income shares with measures of household or familiy inequality. He finds a strong positive relationship, but concludes (P.600): top income shares are far from perfect as a measure of distribution of income across society. Top income shares hence inform not completely on how inequality evolves elsewhere in the distribution. Furthermore, top income shares only weakly satisfy the Pigou-Dalton transfer principle (in contrast to the Gini-Coefficent see above). A transfer from rich to poor will indeed never increase the top income shares, but if the transfer is between invidviduals, who are either both within the top group or both outsie the top group, then the share measure will remain unchanged. \\

@@

An other critical issue with tax data is the problem of tax evasion, which definitely can bias the assessment of inequality. Alvaredo and Saez (2006) for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion. Evasion can occured, when individuals try not to fill tax returns or by misreporting of incomes. In Switzerland non-fillers show up in the tax-statistics either way, as long as they are registered. This person get an imputed income form an older tax return and infomration given by employers. Only non registered non-fillers are not in the records. Therefore non-fillers are a minor problem. Not neglectable is the circumstance, that individuals misreport incomes. \citet{feld_tax_2006} examine the role of tax evasion in Switzerland by calculating the difference of the national accounts measures of primary income and the income reported to the tax authorities. They can show, that the average level of income tax evasion from 1965 to 1995 varies between 13\% and 35\%. They suggest, that evasion is heavily driven by capital income tax evasion. \\

@@

%[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
% Mögliche Vergleichsdimensionen
%- Suitability of cross country comparison
%- Coverage of the population of interesst
%- Implementation of common concepts of economic ressources
%- intertemporal comparision
%- flexibel applications of appropriate inequality measures
% - appropriate income definition

@@

%%
% Population coverage
% Analysen zu den Nullern > ist eine Hauptkritik an der Nutzung von Steuerdaten
% Die Bedeutung von Normal- und Sonderfällen
% Den Vergleich mit der HABE beschreiben, welche Einkommensgrösse nutzen wir aus der HABE um diese möglichst mit den Steuerdaten vergleichbar zu machen? Wir brauchen das Bruttoeinkommen nach Abzug der Sozialversicherungsbeiträge, dies lässt sich mit dem Reineinkommen vergleichen

### 2014-06-13 paper/subdocuments/results.Rnw
minor changes to data and methods and basic structure of results section

\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
Kantonsweise Grafiken
Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}

\subsection{Gini coefficients for taxable income, income before deductions and taxable income after tax}

@@

\subsection{Zero share}

Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later)

\subsection{Income measures with and without zeros}

@@

%Switzerland (income and wealth)
%- graph with gini over time> identify periods of change
%- closer look at interesting periods with relative distribution methods-> is change because of down- or upgrading?
%Cantonal level (income and wealth)
%- different development on regional level 
%Switzerland in international comparison
%- Comparison with other data sources- LIS (inequality in Switzerland is decreasing, which is quite special, because the common pattern in western countries shows an increase in inequality

%- Comparison with other countries (similar data)

% Es scheinen mir aktuell etwas viele Ergebnisse zu sein und ich frage mich, ob wir Ergebnisse auf der Kantonsebene ausklammern sollen. Wenn wir drei Ergebniss haben (1) Unterscheidung von unterschiedlichen Perioden der Ungleichheitsentwicklung (2) Beurteilung dieser Perioden hinsichtlich dem Bereich der Veränderungen (Arme, Reiceh) und Entwicklung der Vermögensungleichheit, dann haben wir bereits viel. Falls wir spezifische Erkenntnisse nur aus dem Kantonsvergleich bekommen (Bias durch Nuller o.ä.), dann sollten wir trotzdem damit arbeiten.

%rf: genau. kantone unterstreichen die probleme mit dem 0er bias und die probleme mit "imputing the gap". mehr muss erstmal nicht sein.

>>>>

% Brauchen wir glaubs nicht mehr
%\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
%Kantonsweise Grafiken
%Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}

% Der Übersicht halber habe ich hier die Titel hier analog zum "Struktur-Issue" auf git gesetzt

\subsection{Defining income}

\emph{Gini coefficients for taxable income, income before deductions and taxable income after tax}

@@

\emph{Gini coefficients considering mean-tested benefits}

\subsection{Measuring inequality}

% Zero-shares (alter Textbaustein)
%Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later) 
% Das allenfalls in der Methodensection machen?
% Vergleich mit der Dell-Reihe zu Non-fillers
% Überschneidungen gib es leider nicht 
% Interessant wäre es auch, zu sehen ob es Brüche in den Zeitreihen der Zahl der Steuerpflichtigen und jener ohne Besteuerung gibt.

\subsection{Population coverage}

\emph{Income measures with and without zeros}

@@

\emph{normal and special cases}

\emph{comparison of tax-data and survey data distribution}

\subsection{Interremporal comparison}
